Validation of Visual Statistical Inference, with Application to Linear Models

### 2012-02-02 paper_jasa.tex
updated reference
updated

% Di: This is specific for the regression model
%Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
Now consider estimating the power of the visual test. For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$. The distribution of $p_B$ is a non-central $t$ distribution under $H_1$ and uniform under $H_0$. 
In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$. Suppose  $p_0=\min_i(p_{0,i})$. Thus $p_0 \sim \text{Beta}(1,m-1)$.

@@

\begin{equation}\label{power_exp} 
   Power(\beta)=Pr(p_{B} < p_0)  \quad \text{for} \beta \ne 0
\end{equation}

@@

\begin{thm}
 For $m \rightarrow \infty $,  Power($\beta$) in equation \ref{power_exp} approaches to $F_{p_B}(0)$, where $F_{p_B}(.)$ is the distribution function of a non-central $t$ distribution. 
 \end{thm}

@@

Again $p_0 \in [0,1]$ and hence $F{p_0}(t) =0$ for $t \le 0$ and  $F{p_0}(t) =1$ for $t \ge 1$. Thus expected power in equation \ref{power_exp} be
\begin{eqnarray*}
 Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
  &=& 1- \int_{-\infty}^{\infty}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
  &=& 1-  \int_{-\infty}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  = 1-  \int_{0}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  \\
  & \rightarrow &1-  \int_{0}^{\infty}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
\end{eqnarray*}

>>>>

% Di: This is specific for the regression model
%Consider the hypothesis test $H_0: \beta_k=0$ against $H_1: \beta_k \ne 0$ in model \ref{multi}.
Now consider estimating the power of the visual test. For the regression slope parameter $\beta$ suppose the associated $p$-value be $p_B$ and its distribution function be  $F_{p_B}(.)$ . As noted by \cite{Ruppert:2007} $F_{p_B}(.)$ is a right skewed distribution under $H_1$ and uniform under $H_0$.  In a lineup plot we simulate $m-1$ residual data sets from null model where each of these $m-1$ null data sets produces a corresponding $p$-value $p_{0,i}$ and $p_{0,i} \sim \text{Uniform}(0,1)$ for $i = 1, ..., m-1$. Suppose  $p_0=\min_i(p_{0,i})$. Thus $p_0 \sim \text{Beta}(1,m-1)$.

@@

\begin{equation}\label{power_exp} 
   Power(\beta)=Pr(p_{B} < p_0)  \quad \text{for}  \quad \beta \ne 0
\end{equation}

@@

\begin{thm}
 For $m \rightarrow \infty $,  Power($\beta$) in equation \ref{power_exp} approaches to $F_{p_B}(0)$.
\end{thm}

@@

Thus expected power in equation \ref{power_exp} be
\begin{eqnarray*}
 Pr(p_{B} < p_0) &=& 1- Pr( p_0 \le p_{B} ) \\
  &=& 1- \int_{-\infty}^{\infty}{Pr( p_0 \le p_{B} |p_{B} =t) f_{p_{B} }(t)dt } \\
  &=& 1-  \int_{-\infty}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  = 1-  \int_{0}^{\infty}{F{p_0}(t) f_{p_{B} }(t)dt }  \\
  & \rightarrow &1-  \int_{0}^{\infty}{ f_{p_{B} }(t)dt }  = F_{p_B}(0) \quad \text{as} \quad m \rightarrow \infty
\end{eqnarray*}
